---
title: "Homework 4"
author: "Lorenzo Galizia, Vincenzo Genna"
date: "16/12/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, warning=FALSE, eval=TRUE, results='hide'}
library(xtable)
library(haplo.stats)
```

```{r, echo=FALSE, include=FALSE}
load("MB.rda")

#2
indiv = nrow(Y)
snps = ncol(Y)
perc = (100 * length(which(is.na(Y))))/length(Y)

tab = data.frame(indiv, snps, perc)

#3
nPossibleHaplotypes = 2 ^ snps

#4
Geno = cbind(substr(Y[,1],1,1),substr(Y[,1],2,2))

for (i in 2:snps) {
  Geno = cbind(Geno, substr(Y[,i],1,1) ,substr(Y[,i],2,2))
}

Snpnames <- paste("SNP",1:snps,sep="")

HaploRes <- haplo.em(Geno,locus.label=Snpnames,control=haplo.em.control(min.posterior=1e-3))


nHaplo = nrow(HaploRes$haplotype)
haploList = HaploRes$haplotype
indexMax = which(HaploRes$hap.prob == max(HaploRes$hap.prob))
mostCommonHaplotype = HaploRes$haplotype[indexMax,]
```


2. Data set information:  

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Individuals & SNPs & Missing data(\%)\\ 
  \hline
 & 139 &  28 & 39.54 \\ 
   \hline
\end{tabular}
\end{table}


3. $2^{nSNP}$ ($=268435456$) haplotypes can theoretically be found if all the genotypes are different.  

4. 6 haplotypes were found using the EM algorithm:  
    + List:
        
        \begin{tabular}{rllllllllll}
          \hline
         & SNP1 & SNP2 & SNP3 & SNP4 & SNP5 & SNP6 & SNP7 & SNP8 & SNP9 & SNP10 \\ 
          \hline
        1 & C & A & A & C & C & A & G & T & A & G \\ 
          2 & T & A & A & C & C & A & G & C & G & G \\ 
          3 & T & A & A & C & C & A & G & C & G & G \\ 
          4 & T & A & A & C & C & A & G & T & A & G \\ 
          5 & T & A & G & G & C & A & G & T & A & G \\ 
          6 & T & C & A & C & C & A & C & T & A & A \\ 
           \hline
        \end{tabular}
        
        \begin{tabular}{rlllllllll}
          \hline
         & SNP11 & SNP12 & SNP13 & SNP14 & SNP15 & SNP16 & SNP17 & SNP18 & SNP19 \\ 
          \hline
        1 & C & G & C & C & A & A & C & C & C \\ 
          2 & C & A & A & T & G & G & T & C & T \\ 
          3 & C & A & A & T & G & G & T & C & T \\ 
          4 & C & A & A & T & G & A & C & C & T \\ 
          5 & C & G & C & C & A & A & C & C & C \\ 
          6 & T & G & C & C & A & A & C & A & C \\ 
           \hline
        \end{tabular}
        
        
        \begin{tabular}{rlllllllll}
          \hline
         & SNP20 & SNP21 & SNP22 & SNP23 & SNP24 & SNP25 & SNP26 & SNP27 & SNP28 \\ 
          \hline
        1 & C & G & C & C & C & C & A & C & T \\ 
          2 & C & A & T & T & C & C & A & C & T \\ 
          3 & C & G & C & T & C & C & A & C & T \\ 
          4 & C & G & C & T & C & C & A & C & T \\ 
          5 & C & G & C & C & C & C & A & C & T \\ 
          6 & G & G & C & C & C & C & A & C & T \\ 
           \hline
        \end{tabular}
        
    + Most common haplotype:  
    
            
        \begin{tabular}{rllllllllll}
          \hline
         & SNP1 & SNP2 & SNP3 & SNP4 & SNP5 & SNP6 & SNP7 & SNP8 & SNP9 & SNP10 \\ 
          \hline
        6 & T & C & A & C & C & A & C & T & A & A \\ 
           \hline
        \end{tabular}
        
        
        \begin{tabular}{rlllllllll}
          \hline
         & SNP11 & SNP12 & SNP13 & SNP14 & SNP15 & SNP16 & SNP17 & SNP18 & SNP19 \\ 
          \hline
        6 & T & G & C & C & A & A & C & A & C \\ 
           \hline
        \end{tabular}
        
        
        \begin{tabular}{rlllllllll}
          \hline
         & SNP20 & SNP21 & SNP22 & SNP23 & SNP24 & SNP25 & SNP26 & SNP27 & SNP28 \\ 
          \hline
        6 & G & G & C & C & C & C & A & C & T \\ 
           \hline
        \end{tabular}  

\newpage    
5. The haplotypic constitution is ambiguous for the individuals in the listIDs: 
    ```{r, echo=FALSE, include=TRUE}
    #5
    #names(table(Y[4,]))[2]
    #substr(names(table(Y[4,]))[2],1,1) != substr(names(table(Y[4,]))[2],2,2)
    
    listIds=c()
    for(i in 1:nrow(Y)) {
      cont=0
      for(j in 1:length(table(Y[i,]))) {
        if(substr(names(table(Y[i,]))[j],1,1) != substr(names(table(Y[i,]))[j],2,2)) {
          cont=cont+table(Y[i,])[names(table(Y[i,]))[j]]
        }
      }
      if(cont>1) {
        listIds[length(listIds)+1]=i;
      }
    }
    listIds
    ```

6. Sembra che i risultati non cambiano però il bisognerebe sapere il perchè...
    + List:
        
        \begin{tabular}{rllllllllll}
          \hline
         & SNP1 & SNP2 & SNP3 & SNP4 & SNP5 & SNP6 & SNP7 & SNP8 & SNP9 & SNP10 \\ 
          \hline
        1 & C & A & A & C & A & G & T & A & G & C \\ 
          2 & T & A & A & C & A & G & C & G & G & C \\ 
          3 & T & A & A & C & A & G & C & G & G & C \\ 
          4 & T & A & A & C & A & G & T & A & G & C \\ 
          5 & T & A & G & G & A & G & T & A & G & C \\ 
          6 & T & C & A & C & A & C & T & A & A & T \\ 
           \hline
        \end{tabular}
        
        \begin{tabular}{rlllllllll}
          \hline
         & SNP11 & SNP12 & SNP13 & SNP14 & SNP15 & SNP16 & SNP17 & SNP18 & SNP19 \\ 
          \hline
        1 & G & C & C & A & A & C & C & C & C \\ 
          2 & A & A & T & G & G & T & C & T & C \\ 
          3 & A & A & T & G & G & T & C & T & C \\ 
          4 & A & A & T & G & A & C & C & T & C \\ 
          5 & G & C & C & A & A & C & C & C & C \\ 
          6 & G & C & C & A & A & C & A & C & G \\ 
           \hline
        \end{tabular}
        
        
        \begin{tabular}{rllllllll}
          \hline
         & SNP20 & SNP21 & SNP22 & SNP23 & SNP24 & SNP25 & SNP26 & SNP27 \\ 
          \hline
        1 & G & C & C & C & C & A & C & T \\ 
          2 & A & T & T & C & C & A & C & T \\ 
          3 & G & C & T & C & C & A & C & T \\ 
          4 & G & C & T & C & C & A & C & T \\ 
          5 & G & C & C & C & C & A & C & T \\ 
          6 & G & C & C & C & C & A & C & T \\ 
           \hline
        \end{tabular}
        
    + Haplotype frequencies:  
    
        \begin{table}[ht]
        \centering
        \begin{tabular}{rrrrrrr}
          \hline
         & 1 & 2 & 3 & 4 & 5 & 6 \\ 
          \hline
        freq & 0.02 & 0.06 & 0.02 & 0.00 & 0.18 & 0.73 \\ 
           \hline
        \end{tabular}
        \end{table}
